---
title: "Basics of magrittr in R"
output:
  html_document:
    toc: true
---

```{r, echo=FALSE,fig.height=4,fig.width=6,fig.align='center'}
suppressWarnings(suppressPackageStartupMessages(library(png)))
suppressPackageStartupMessages(library(grid))
img <- readPNG('magrittr.png')
grid.raster(img)
```

The [magrittr](https://github.com/smbache/magrittr) package offers a new operator that can help improve readability of your code, and make it easier to update and modify data wrangling code.  The %>% operator has been adopted into `dplyr` and many of [Hadley Wickham's](https://github.com/hadley) packages are written to be pipe-friendly.

# The Problem

### R code can get hard to read

```{r, eval=FALSE}
sapply(iris[iris$Sepal.Length < mean(iris$Sepal.Length),-5],FUN = mean)
```

### A (Possible) Solution - the pipe %>% {.build}

- Similar to Unix pipe |
- Code can be written in the order of execution, left to right
- `%>%` will "pipe" information from one statement to the next
  * `x %>% f` is equivalent to `f(x)`
  * `x %>% f(y)` is equivalent to `f(x,y)`
  * `x %>% f %>% g %>% h` is equivalent to `h(g(f(x)))`

<b>{magittr} provides 4 special operators</b>

- `%>%` - pipe operators
- `%T>%` - tee operator
- `%$%` - exposition operator
- `%<>%` - compound assignment pipe operator

```{r, echo=FALSE, warning=FALSE,message=FALSE}
library(magrittr,quietly=T)
library(dplyr,quietly=T)
```

<a href="#top">Back to top</a>

# What %>% is doing

The `%>%` is taking the output of the left-hand side and using that for
the first argument of the right-hand side, or where it finds a *`.`*

### Basic Example

```{r}
df <- data.frame(x1=rnorm(100),x2=rnorm(100),x3=rnorm(100))

df %>% head(1)  # same as using head(df,1)
df %>% head(.,1)  # same as using head(df,1)
```

### Slightly more complicated example

```{r,fig.height=5,fig.width=5,fig.align='center'}
library(ggplot2)
mtcars %>%
  xtabs(~gear+carb,data=.) %>%
  as.data.frame %>%
  ggplot(.,aes(x=gear,y=carb,size=Freq)) +
  geom_point()
```






### Even more complicated example

```{r, eval=FALSE}
# Generate some sample data.
df <-
  data.frame(
    Price    = 1:100 %>% sample(replace = TRUE),
    Quantity = 1:10  %>% sample(replace = TRUE),
    Type     =
      0:1 %>%
      sample(replace = TRUE) %>%
      factor(labels = c("Buy", "Sell"))
  )
```
[Source](http://www.r-statistics.com/2014/08/simpler-r-coding-with-pipes-the-present-and-future-of-the-magrittr-package/)


<a href="#top">Back to top</a>


```{r,eval=TRUE,echo=FALSE,message=FALSE,warning=FALSE}
# This is because the stats package has a filter function also, and ggplot2 depends on stats
unloadNamespace('dplyr')
library(dplyr)
```

## The combination of %>% with {dplyr}

- `filter()`
- `group_by()`
- `summarise()`,`summarize()`
- `arrange()`
- `mutate()`
- `select()`


```{r}
sapply(iris[iris$Sepal.Length < mean(iris$Sepal.Length),-5],FUN = mean)
iris %>%
  mutate(avg.length=mean(Sepal.Length)) %>%
  filter(Sepal.Length<avg.length) %>%
  select(-Species,-avg.length) %>%
  summarise_each(funs(mean))
```

<a href="#top">Back to top</a>

## %$% The exposition operator

- Similar to `with()` or `attach()`
- Useful for functions that don't take a data parameter

```{r}
table(CO2$Treatment,CO2$Type)
# with(CO2,table(Treatment,Type))
CO2 %$% table(Treatment,Type)
```

<a href="#top">Back to top</a>

## %T>% The Tee Operator

- Allows a "break" in the pipe.
- Executes right-hand side of `%T>%`, but will continue to pipe through to next statement

```{r}
iris %>%
filter(Species != 'virginica') %>%
select(Sepal.Width,Sepal.Length) %T>%
plot %>%  # Make scatterplot and keep going
colMeans
```

<a href="#top">Back to top</a>

## %<>% The Compound Assignment Operator
- Combines a pipe and an assignment operator
- Think `i++` or `x+=z` from the C family, Python, Ruby, etc.

<a href="#top">Back to top</a>

## %<>% The Compound Assignment Operator
```{r}
df <- rexp(5,.5) %>% data.frame(col1=.)
df
df %<>% arrange(col1)
df
```

<a href="#top">Back to top</a>

# More Resources
- [Source as RMarkdown](https://github.com/rweyant/bertplot/blob/master/R/tutorials/magrittr-basics/magrittr-basics.Rmd)
- [magrittr on GitHub](https://github.com/smbache/magrittr)
- [dplyr and pipes: the basics](http://seananderson.ca/2014/09/13/dplyr-intro.html)
- [pipeR vs. magrittr](http://renkun.me/blog/2014/08/08/difference-between-magrittr-and-pipeR.html)
